N-Body problem

### 2014-10-25 article/article.tex
typos, close encounter sectie

To verify the accuracy of the simulation we can apply some fundamental conservation laws from physics to our simulation. We know based on empirical data that these laws seem to be true for the natural, so if our simulation is a faithful model of nature these laws should also be true for our simulation. If they do not, the amount our simulation fails to conserve certain values that are supposed to be conserved can be used as a metric for the error in our simulation.\\\\

@@

The total (mechanical) energy of system is the potential and kinetic energy of the system added together, $E = W+K$. This total energy should remain constant. In a simulation this can however go wrong. For instance a body may experience a certain strong potential energy. In reality it would only experience this energy for a very short time. But in a simulation with discrete timesteps, the timestep may be longer than the time the body is supposed to experience the potential energy for. Thus it will create extra kinetic energy without that energy being based on potential energy that would be there in reality when it interact continuously with bodies rather than at discrete timesteps. \\

@@

We have opted to go for a direct simulator, because the implementation of such algorithms tends to be less complex. The important part in these simulations is the determination of the force each of the bodies experiences. In direct solvers, this is done the same for each body to body interactions and the total force on a particular body is a simple summation of the forces it experiences from every other body. \\
To determine the force between two bodies due to gravity is a second order differential equation needs to be solved. Thus numerical integrators are the key component of direct solvers.

@@

>>>>

To verify the accuracy of the simulation we can apply some fundamental conservation laws from physics to our simulation. We know based on empirical data that these laws seem to be true for the natural world, so if our simulation is a faithful model of nature these laws should also be true for our simulation. If they do not, the amount our simulation fails to conserve certain values that are supposed to be conserved can be used as a metric for the error in our simulation.\\\\

@@

The total (mechanical) energy of a system is the potential and kinetic energy of the system added together, $E = W+K$. This total energy should remain constant. In a simulation this can however go wrong. For instance a body may experience a certain strong potential energy. In reality it would only experience this energy for a very short time. But in a simulation with discrete timesteps, the timestep may be longer than the time the body is supposed to experience the potential energy for. Thus it will create extra kinetic energy without that energy being based on potential energy that would be there in reality when it interact continuously with bodies rather than at discrete timesteps. \\

@@

We have opted to go for a direct simulator, because the implementation of such algorithms tends to be less complex. The important part in these simulations is the determination of the force each of the bodies experiences. In direct solvers, this is done the same for each body to body interaction and the total force on a particular body is a simple summation of the forces it experiences from every other body. \\
To determine the force between two bodies due to gravity a second order differential equation needs to be solved. Thus numerical integrators are the key component of direct solvers.

@@

\textbf{Handling Close Encounters (not of the Third Kind)}\\
In a gravitational simulation with discrete timesteps strange things can happen if two bodies get really close to each other. If two bodies are really close to each other they may experience a very strong force interaction. In reality this would happen for a very short time before the bodies either collide or they move away from each other again and experience a decreased amount of force.\\
The n-body simulation that has been studied uses a collisionless n-body model. This means that the first option for close encounters (collision) is disregarded. A simulation with discrete timesteps will assume that the force experienced at the start of a timestep by a body will be experienced by that body during the whole timestep. This is not necessarily true for two close bodies in collisionless system. The bodies would only experience the strong force for a much shorter time than the timestep actually last and thus the force interaction gets overestimated. \\
One option to improve accuracy of close encounters is to have the simulation run with smaller timesteps. This would however increase computational load and be inefficient if most simulation steps don't actually have encounters close enough to need a shorter timestep.\\
Instead a dynamic timestep system can be devised. By looking at each gravitational interaction in the system, the resulting speeds and accelerations between every pair of bodies due to force interaction can be derived. It can then be calculated how long it would take to cover the distance between the two bodies based on their force interaction. If the length of time is shorter than the default timestep length, the timestep for the simulation is lowered for the next simulation step.\\ 
This lowering of the timestep is done globally, for the entire simulation. The simulation is slowed down based on the closest interaction between two bodies in the system. There are algorithms that lower the timestep per body, but those have not been looked at.
